Estimating SARS-CoV-2 transmission in England from randomly sampled prevalence surveys

### 2023-12-21 transmission.tex
minor edits

Throughout this chapter, the CIS data will be stratified into the regions of England and the following age groups.
\begin{enumerate}
    \item $[0, 11)$: children of pre-school and primary school age.
    \item $[11, 16)$: children of compulsory secondary-school age.
    \item $[16, 25)$: adolescents and young adults who may remain in formal education.
    \item $[25, 50)$: younger working age adults.
    \item $[50, 70)$: older working age adults.
    \item $[70, \infty)$: elderly.
\end{enumerate}
These groupings align closely with those previously used for analysis of the CIS data~\autocites[e.g.:][]{houseInferring}{walkerTracking}.
\todo{Contrast with age groupings used in previous chapter}

@@

@@

Some behavioural changes (\eg mask wearing) impact transmission.
However, these are not captured by the contact survey.
To allow for these changes and the uncertainty produced in formulating $\matr{\kaapa}(t)$, I follow \textcite{birrellRealtime} in placing a random walk prior on $\log\beta(t)$  on the weekly timescale:
\begin{align}
    \beta(t) &= \beta_{\tilde{w}(t)} \\
    \log\beta_{w+1} &= \log\beta_w + \sigma_\epsilon \epsilon_w \\
    \epsilon_w &\sim \N(0, 1)
\end{align}
where $\tilde{w}(t)$ is the week number of time $t$, with week 1 being the first week considered.
Two hyperparameters will be inferred: $\beta_1$, the initial value of the random walk; and $\sigma_\epsilon$, the standard deviation of the random walk, will have priors placed on them (see \cref{SEIR:sec:priors}).

@@

The majority of the confidence intervals contain the nominal probability (see \cref{SEIR:fig:sim-coverage}).
There is a tendency for the coverage to be slightly lower than the nominal probability, suggesting that the credible intervals are slightly too narrow; this is a common problem when using random-walk MCMC.
\todo{cite issues with random-walk MCMC exploring space}
This is similar when considering coverage of the derived quantities incidence and prevalence (see \cref{SEIR:fig:sim-inc-prev}).
\begin{figure}
    \includegraphics{SEIR/sim/coverage}
    \caption[Coverage of simulation study]{%
        Point estimate and 95\% confidence interval for coverage of the 50\%, 75\%, and 95\% central credible intervals for each parameter; see text for full definition.
        Vertical lines show the nominal coverage.
    }
    \label{SEIR:fig:sim-coverage}
\end{figure}
\begin{figure}
    \thisfloatpagestyle{empty}
    \vspace{-3cm}
    \makebox[\textwidth][c]{\includegraphics[width=0.9\paperwidth]{SEIR/sim/predictive_coverage}}
    \caption[Coverage of simulation study (derived quantities)]{%
        Point estimate and 95\% confidence interval for coverage of the 50\%, 75\%, and 95\% central credible intervals for incidence (top) and prevalence (bottom); see text for full definition.
        Horizontal lines show the nominal coverage.
    }
    \label{SEIR:fig:sim-inc-prev}
\end{figure}

\todo[inline]{What happened to the following paragraph? It doesn't make sense currently!}
In this setup, the parameters $i^+$, $q_c$, and $\psi$ are fully recovered; the other parameters are only partially so: their posterior has moved away from the prior but is still regularised by it (see \cref{SEIR:fig:true-vs-posterior}).
The ability to identify $q_c$ is an advantage of using data which includes measurements on children, unlike data on hospitalisations or deaths.
This suggests that these parameters are not practically identifiable in this setting.
The posteriors align very closely with the priors (not shown) which reinforces this finding.
The lack of identifiability for $\vec{\pi}$ aligns with previous work showing this parameter is structurally unidentifiable, \ie not identifiable even in the limit of infinite data~\autocite{dankwaStructural}.
\begin{figure}
    \includegraphics{SEIR/sim/true_vs_posterior}
    \caption[True vs posterior parameter values]{%
        True vs posterior median parameter values for each parameter in each simulation.
        The black line shows the line $y = x$ which the dots would lie on if the parameter is fully recovered.
        The blue line is a linear regression line.
        For $i^+$, $q_c$, and $\psi$, the posterior recovers the true values.
        For $\vec{\pi}$ and $\rho$, the posterior median is very similar regardless of the true value.
    }
    \label{SEIR:fig:true-vs-posterior}
\end{figure}

>>>>

Throughout this chapter, the CIS data will be stratified into the regions of England and the following age groups.
\begin{enumerate}
    \item $[0, 11)$: children of pre-school and primary school age.
    \item $[11, 16)$: children of compulsory secondary-school age.
    \item $[16, 25)$: adolescents and young adults who may remain in formal education.
    \item $[25, 50)$: younger working age adults.
    \item $[50, 70)$: older working age adults.
    \item $[70, \infty)$: elderly.
\end{enumerate}
These groupings align closely with those previously used for analysis of the CIS data~\autocites[e.g.:][]{houseInferring}{walkerTracking}.

@@

This section uses the same notation as \cref{E-SEIR:sec:mechanistic-models}, following the infectious disease modelling literature rather than the statistical conventions used elsewhere in this thesis.

@@

Some behavioural changes (\eg mask wearing) impact transmission.
However, these are not captured by the contact survey.
To allow for these changes and the uncertainty produced in formulating $\matr{\kappa}(t)$, I follow \textcite{birrellRealtime} in placing a random walk prior on $\log\beta(t)$  on the weekly timescale:
\begin{align}
    \beta(t) &= \beta_{\tilde{w}(t)} \\
    \log\beta_{w+1} &= \log\beta_w + \sigma_\epsilon \epsilon_w \\
    \epsilon_w &\sim \N(0, 1)
\end{align}
where $\tilde{w}(t)$ is the week number of time $t$, with week 1 being the first week considered.
Two hyperparameters will be inferred: $\beta_1$, the initial value of the random walk; and $\sigma_\epsilon$, the standard deviation of the random walk, will have priors placed on them (see \cref{SEIR:sec:priors}).

@@

The majority of the confidence intervals contain the nominal probability (see \cref{SEIR:fig:sim-coverage}).
There is a tendency for the coverage to be slightly lower than the nominal probability, suggesting that the credible intervals are slightly too narrow.
It is well-known that random walk MCMC, as used here, can poorly explore the tails of the posterior in finite time~\autocite[e.g.][]{turitsynIrreversible}.
\todo{cite issues with random-walk MCMC exploring space}
This is similar when considering coverage of the derived quantities incidence and prevalence (see \cref{SEIR:fig:sim-inc-prev}).
\begin{figure}
    \includegraphics{SEIR/sim/coverage}
    \caption[Coverage of simulation study]{%
        Point estimate and 95\% confidence interval for coverage of the 50\%, 75\%, and 95\% central credible intervals for each parameter; see text for full definition.
        Vertical lines show the nominal coverage.
    }
    \label{SEIR:fig:sim-coverage}
\end{figure}
\begin{figure}
    \thisfloatpagestyle{empty}
    \vspace{-3cm}
    \makebox[\textwidth][c]{\includegraphics[width=0.9\paperwidth]{SEIR/sim/predictive_coverage}}
    \caption[Coverage of simulation study (derived quantities)]{%
        Point estimate and 95\% confidence interval for coverage of the 50\%, 75\%, and 95\% central credible intervals for incidence (top) and prevalence (bottom); see text for full definition.
        Horizontal lines show the nominal coverage.
    }
    \label{SEIR:fig:sim-inc-prev}
\end{figure}

\todo[inline]{What happened to the following paragraph? It doesn't make sense currently!}
In this setup, the parameters $i^+$, $q_c$, and $\psi$ are fully recovered; the other parameters are only partially so: their posterior has moved away from the prior but is still regularised by it (see \cref{SEIR:fig:true-vs-posterior}).
The ability to identify $q_c$ is an advantage of using data which includes measurements on children, unlike data on hospitalisations or deaths.
This suggests that these parameters are not practically identifiable in this setting.
The posteriors align very closely with the priors (not shown\todo{consider showing prior/posterior plots}) which reinforces this finding.
The lack of identifiability for $\vec{\pi}$ aligns with previous work showing this parameter is structurally unidentifiable, \ie not identifiable even in the limit of infinite data~\autocite{dankwaStructural}.
\begin{figure}
    \makebox[\textwidth][c]{\includegraphics{SEIR/sim/true_vs_posterior}}
    \caption[True vs posterior parameter values]{%
        True vs posterior median parameter values for each parameter in each simulation.
        The black line shows the line $y = x$ which the dots would lie on if the parameter is fully recovered.
        The blue line is a linear regression line.
        For $i^+$, $q_c$, and $\psi$, the posterior recovers the true values.
        For $\vec{\pi}$ and $\rho$, the posterior median is very similar regardless of the true value.
    }
    \label{SEIR:fig:true-vs-posterior}
\end{figure}
